### 2021-06-11 abstract.tex

\chapter*{Abstract}

There is a growing interest in shape analysis in recent years and we present a novel contour-based shape representation for 2D bounded simply-connected domains in this thesis, named the Harmonic Beltrami signature. The proposed signature is based on the harmonic extension of the conformal welding map of a unit circle. With suitable normalization, the uniqueness of this harmonic function is guaranteed up to rotations. And finally quasi-conformal theory helps get rid of the only uncertainty by computing the Beltrami coefficient of harmonic extension. There exists one-by-one relationship between shapes and our Harmonic Beltrami signatures and we also give the method to reconstruct the original shapes from our signatures. The benifit of the proposed signature is that it keeps to be invariant under simple transformations like sacling, transformation and rotation and is roubost under slight deformations and distortions. Experiments demonstrate the above properties and also show the excellent classification performance.