### 2021-06-11 acknowledgement.tex

I am deeply grateful to Prof. Ronald Lui not only for his guidance in my two-year M.Phil. study, but also bringing me to the world of applied mathematics and computational differential geometry. The opportunity and support he has given to me is unprecedented, which cultivates me to be an independent researcher at the same time.

Besides, I would also like to thank Dr. ZHANG Daoping for his help and discussions in my projects. His advice really helped me overcome obstacles and brought insights and convenience to me.

I would also like to thank my family for their consistent love, support and understanding to me, without which I can never persist on my way studying and doing mathematics.

Besides, I am also grateful to my friends Mr. LI Zhiwen, Mr. CHEN Qiguang, Mr. LIU Oscar, to name a few, for the happy moments we have in my daily life.

Moreover, I am grateful to CUHK and the math department, for providing support and resources on my way towards the master's degree.

Lastly, thank you my wife Yanyi, for your consistent love and support.